### 2013-08-02 itm/appendix.tex
copy *& paste everything

\documentclass[runningheads]{llncs}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage[numbers]{natbib}
\usepackage{csquotes}
\usepackage[natbib, style=numeric-comp, firstinits=true, sorting=none, maxnames=2]{biblatex}

\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
\mainmatter

\begin{document}

\begin{theorem}
Consider the two cluster case with $n$ data points.
Let $n_0, n_1$ be the cluster sizes and $L_0, L_1$ be the length of the minimum spanning trees in the obtimum solution.
The non-optimality of a solution obtained by ITM the search as above is bounded as
\begin{align}
    \hat{I}^* - \hat{I}_{\text{ITM}} \leq 
         d \left [ \log \Bigl(\frac{L_0 + e}{L_0}\Bigr) + \frac{n_1}{n} \log \Bigl(\frac{L_0}{L_1} \Bigr) \right] 
\end{align}
\begin{proof}
Remove an edge of the MST with such that the resulting clustering with
MST lengths $L'_0, L'_1$ and $n'_0, n'_1$ the corresponding cluster sizes fulfill $L'_0 \geq L'_1$ and $n'_1 \geq n'_0$.

Wlog. $L_0 \geq L_1$.
Let $2e$ be the length of the longest edge in the MST of the whole dataset minus its shortest edge.
Then, by construction $2 (L_0 + e) \geq L_0 + L_1 + 2e \geq L'_0 + L'_1$.
By monotonicity of the logarithm, $\log(L_0 + e) \geq \log(\frac{L'_0 + L'_1}{2})$.
Using the concavity of the logarithm, we obtain
\begin{align}\label{concavity}
    \log(L_0 + e) \geq \frac{1}{2} \log(L'_0) + \frac{1}{2} \log(L'_1) \geq \frac{n_0}{n}\log(L'_0) + \frac{n_1}{n}\log(L'_1)
\end{align}

Any clustering obtained in this way is certain to have an objective not higher than the solution obtained by ITM\@.
For ease of analysis we will compare the optimum solution against any clustering that fulfills this restriction.
Then 
\begin{align}
    \hat{I}^* - \hat{I}_{\text{ITM}}
    & = -\frac{d}{n} \left [n_0 \log(L_0) + n_1 \log(L_1) - n_0 \log(L'_0) - n_1 \log(L'_1) \right ] + Z\\
    %& = -\frac{d}{n} \left [n_0 \log(L_0) + n_1 \log(L_0) + n_1 \log \Bigl (\frac{L_1}{L_0} \Bigr ) - n_0 \log(L'_0) - n_1 \log(L'_1) \right ] \\
    & = -\frac{d}{n} \left [n \log(L_0) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) - n_0 \log(L'_0) - n_1 \log(L'_1) \right] +Z\\
    & \leq -\frac{d}{n} \left [n \log(L_0) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) - n\log(L_0 + e) \right] +Z\\
    & = -\frac{d}{n} \left [n \log(\frac{L_0}{L_0 + e}) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) \right] +Z\\
    & = d \left [ \log(\frac{L_0 + e}{L_0}) + \frac{n_1}{n} \log \Bigl(\frac{L_0}{L_1} \Bigr) \right] + Z, 
\end{align}
which concludes the proof.
\end{proof}

\end{theorem}
\end{document}