### 2013-09-15 nyu/legend.tex
add icra nyu paper

\documentclass{standalone}
\usepackage{color}
\usepackage{tikz}

\definecolor{structure}{RGB}{185, 202, 202}
\definecolor{prop}{RGB}{227, 172, 0}
\definecolor{furniture}{RGB}{255, 94, 28}
\definecolor{ground}{RGB}{9, 68, 83}
%[``structure'', ``prop'', ``furniture'', ``ground'' , ``void'']

\begin{document}
    \begin{tikzpicture}
        \matrix [column sep=.5cm]{%
            \node [rectangle, fill=ground, minimum width=.5cm,minimum height=.5cm] {};&
        \node [rectangle] {Ground};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=structure, minimum width=.5cm,minimum height=.5cm] {};&
        \node [rectangle] {Structure};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=furniture, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Furniture};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=prop, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Props};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=black, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Void};\\
    };
    \end{tikzpicture}
\end{document}