### 2013-09-19 intro.tex
say everything twice. comit and delete lol

\chapter{Introduction}
\vspace{1cm}
\hfill%
\begin{minipage}{.4\linewidth}
    %\[ \argmax_{y \in \mathcal{Y}} f(x, y) \]
    %\dots I wish I could let it stand on its own like this, but of course there is more to come.
    %\flushright
    %\emph{Martial Hebert}
%Remember that all models are wrong;\\
%the practical question is how wrong do they have to be to not be useful.
Essentially, all models are wrong, but some are useful. 
\flushright%
\emph{George E. P. Box}

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lack behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes, and only have limited capabilities to model interactions or relations.
We believe that context is one of the most important cues when it comes to classifying objects,
and therefore understanding scenes. Consequently, we target dense labeling of scenes on a fine-grained level.
The task of densely labeling image pixels into classes is called object class segmentation
or semantic segmentation.
\pagebreak

We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of \emph{context}.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will emerge soon.

We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent and process.

Before we delve into semantic segmentation and object class recognition, we
will first investigate general grouping mechanisms in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. Chapter~\ref{ch:itm} introduces a novel, information theoretic
algorithm that compares favourably with algorithms from the literature.
While we do not apply our algorithm to the task of bottom-up segmentation,
this is a promising avenue for future research.

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is only available through laborious manual annotation. 
Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
shortcoming of object class segmentation approaches by introducing a method to
find segmentations automatically from image level annotations.

The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We will focus on the use of conditional random fields~(CRFs), which have shown
great promise for computer vision applications. Using the paradigm of
structural support vector machines, it is possible to learn conditional random
fields to directly minimize a loss of interest. In particular, conditional
random fields allow to combine different cues, possibly produced by other
approaches, in a principled manner. One of the main difficulties with CRF
approaches to computer vision problems is, that context in images is usually
represented as a neighborhood graph.  These graphs, by nature, contain many
cycles, making inference intractable in general. Consequently, learning algorithms
have to rely on approximate inference, with often unknown consequences to learning.

There have been several previous studies on learning structural support vector machines,
and learning for conditional random fields. The impact of approximate inference was first
investigated by \citet{finley2008training}, applying structural support vector machines
to multi-label data. Later, different works investigated how to combine approximate inference and
learning in a single framework. \citet{meshi2010learning, komodakis2011efficient}, and \citet{hazan2010primal} approached
the problem using duality, and formulate learning and inference as a joint optimization problem.
\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and  \citet{krahenbuhlparameter} formulate learning
structured prediction as optimizing an \emph{prediction engine}, that takes into account all aspects
of the model, in particular the inference algorithm used.
In this work, on the other hand, we follow the well-established algorithms for learning structural
support vector machines, and investigate how we can use the available inference algorithms to
obtain good results within a reasonable time-frame.

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where inference is simple.
More recently, \citet{lucchi2013learning} proposed a novel algorithm for
efficiently learning structured prediction for semantic segmentation, using
approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
algorithm that runs in similar time to the one proposed by
\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
the same dataset.

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introduction of general clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstration of an algorithm for  weakly supervised object class segmentation.
\item Providing a general efficient implementation for structured prediction.
\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
%\item Shows importance of edge features.
\end{itemize}

\section{Thesis Outline}
We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
which we plan to use as an additional building block for semantic segmentation
algorithms in future work. Clustering is important both for the initial
segmentation into superpixels, and in the creation of visual words.

We will then introducing a mostly unstructured
approach to semi-supervised learning for object class segmentation in
Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation. The central topic of this thesis is
introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
software library for implementing structured learning and prediction
algorithms.
Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

The problem of learning with approximate inference is investigated in
Chapter~\ref{ch:exact_learning}.  We develop a strategy for efficient caching
and a combination of inference algorithms that allows us to learn SSVMs for
semantic image segmentation exactly, even though the involved factor graphs
contain many loops. We demonstrate our algorithm on the Pascal VOC 2010 and MSRC-21 datasets,
and provide state-of-the-art results on the MSRC-21 dataset.

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structured_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spacial relations, and improve
upon the state-of-the-art in the NYU V2 datasets that we consider.

\section{Publications}
The main material of this thesis has either been published in conference
proceedings or has been submitted to conferences or journals. We now list the
relevant publications.
\begin{description}
    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and Christoph H. Lampert. Published in the proceedings of the German Conference on Pattern Recognition.
    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven Behnke. Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of Machine Learning Research, Open Source Software track.
    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Computer Vision Theory and Applications.
    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Robotics and Automation.
\end{description}