### 2013-09-25 evaluation/paper.tex
more results, discussion for experiments

\chapter{Empirical Comparison of Learning Algorithms}\label{ch:comparison}
In this chapter, we provide an empirical evaluation of the learning algorithms
described in Chapter~\ref{ch:structured_pystruct}.  We use the open source
implementations in \pystruct, and publish the evaluation code and datasets with
the package.

\section{Datasets and Models}
We consider several qualitatively different datasets, that have been widely used in the literature.
The problems we consider are multi-class classification, sequence labeling, multi-label prediction,
and general graph labeling.

\subsection{Multi-Class Classification (MNIST)}
\begin{figure}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_0}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_1}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_2}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_3}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_4}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_5}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_6}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_7}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_8}
    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_9}
    \caption{%
        Visualization of samples from the ten classes in the MNIST dataset.\figlabel{mnist_illustration}
    }
\end{figure}
The simplest task we use is multi-class classification. In this task,
the inference problem is trivial, as it is assumed the target set is small enough
to be enumerated efficiently. While this problem could be solved more efficiently
with specialized algorithms, it nevertheless provides an initial insight into structured
prediction algorithms.
We choose the classical MNIST dataset of handwritten digits, consisting of
60000 training images and 10000 test images of the digits zero to nine.
\Figref{mnist_illustration} shows some examples. Each image is a $28 \times 28$
grey-level image, resulting in a $784$-dimensional feature vector. We normalize
the feature to lie between $0$ and $1$.
%
\enlargethispage{15mm}
The model we use for multi-class classification is the Crammer-Singer
formulation. We do not include a bias, leading to $784 \cdot 10 = 7840$
parameters in the model.

\begin{figure}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_00}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_01}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_02}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_03}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_04}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_05}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_06}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_07}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_08}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_09}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_10}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_11}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_12}\\[3mm]
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_00}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_01}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_02}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_03}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_04}\\[3mm]
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_00}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_01}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_02}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_03}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_04}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_05}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_06}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_07}
    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_08}\\
    \caption{%
        Visualization of some words from the OCR dataset. The first letters were removed by \citet{taskar2003max} as 
        they were capitals.
        The words are: (j)ustifications, (s)kiing and (u)nexpected. The letters in parentheses are not part of the dataset.
        \figlabel{ocr_illustration}
    }
\end{figure}
\subsection{Sequence Labeling (OCR)}
A classical application of structured prediction is sequence labeling.
We choose the ``OCR'' dataset introduced in the seminal work of \citet{taskar2003max}.
Each input consists of the segmented handwritten letters of a word in lower
case. The task is to classify all letters in a word, that is assign each
segmented input letter one of the classes ``a'' to ``z''. As the first letter
of each word was capitalized, these were removed by \citet{taskar2003max},
leading to somewhat odd-looking labels. The words are between three and fourteen
letters long, and each letter is represented as a binary image of size $16
\time 8$, with a total of $6877$ words.
The dataset is divided into ten folds. We consider two setups: learning
on one fold, and testing on the remaining nine folds, following \citet{taskar2003max},
and learning on nine folds and testing on the remaining fold, following \citet{lacoste}.
We refer to the learning on one fold as OCR-small, and learning on nine folds as OCR-big.
We use a simple chain model with a single constant pairwise feature.
This means the unary potential has $16 \cdot 8 \cdot 26 = 768$ parameters, one for each input
feature and output class. The pairwise potential consists of a matrix of
transition potentials with $26 \cdot 26=676$ entries.
It is well-known that efficient exact inference in chains is possible using
message passing algorithms.

\subsection{Multi-Label Classification}
Multi-label classification is a generalization of multi-class classification,
in which each example can be associated with more than one class. In other
words, the algorithm must decide for each sample and for each class whether the
sample belongs to that class or not.
Multi-label classification was first formulated as a structured prediction problem
in by \citet{finley2008training}, where it was used to investigate the influence of
approximate inference on the $n$-slack cutting plane algorithm. In this formulation,
each class is represented as a binary node in a factor graph---the states representing
presence of absence of the class. A different factor of pairwise potentials is introduced
between each pair of classes.
We also consider a different model, where pairwise potentials are only
introduced between specific nodes. We build a tree over the binary variables by
computing the Chow-Liu Tree~\citep{chow1968approximating} of the targets. While
this results in a less expressive model, a tree-shaped model allows for exact
inference via message passing.
We use two of the datasets used in \citet{finley2008training}, the scene
and yeast datasets. We choose these two as these are real-world datasets for
which the pairwise approach outlined above actually improves upon the
baseline~\citep{finley2008training}.
The scene dataset has six labels, 294 input features, 1211 training samples and
1196 test samples. This leads to $204 \cdot 6 = 1224$ parameters for the unary potentials,
$3 \cdot 5 \cdot 4 = 60$ parameters for the full pairwise potentials (four for each edge), and
$5 \cdot 4 = 20$ parameters for the pairwise potentials of the tree-shaped model.
Using four parameters for each edge is a slight over-parametrization that
simplifies writing out the model.
The yeast dataset has 14 labels, 103 features, 1500 training samples and 971
test samples. The resulting numbers of parameters are $14 \cdot 103=1442$ parameters
for the unary potentials, $7 \cdot 13 \cdot 4=364$ parameters for the full pairwise
potential, and $13 \cdot 4=52$ parameters for the tree-shaped model.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XXXXXXX@{}}
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0007}\\
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0007}
    \end{tabu}
\caption{%
    Visualization of the snakes dataset. The top row shows input patterns, the bottom row the corresponding
    labels. The colors in the image showing the labels correspond to dark blue for background, and encoding
    the length of the snake from head (red) to tail (blue).
\figlabel{snakes_illustration}}
\end{figure}

\subsection{2D Grid CRF (Snakes)}
The snakes dataset is a synthetic dataset, where samples are labeled 2D grids.
It was introduced in \citet{nowozin2011decision} to demonstrate the importance
of learning conditional pairwise potentials. The dataset consists of ``snakes''
of length ten traversing the 2D grid. Each grid cell that was visited is marked
as the snake heading out towards the top, bottom, left, or right, while
unvisited cells are marked as background.  The goal is to predict a labeling of
the snake from ``head'' to ``tail'', that is assigning numbers from zero to
nine to the cells that are occupied by the snake.  \Figref{snakes_illustration} illustrates the
principle.
Local evidence for the target label is weak except
around head and tail, making this a challenging task. Pairwise potentials are necessary to solve the
task. The dataset is noise-free in the sense that given the above description, a
human could easily produce the desired labeling without making any mistake.
The dataset is also interesting as the model of \citet{nowozin2011decision} produced
notoriously hard to optimize energy functions.

In \citet{nowozin2011decision}, the input was encoded into five RGB colors
(``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
more suitable for our linear methods, we convert this representation to a
one-hot encoding of the five states.

We use a grid CRF model for this task. Unary for each node are given by the
input of the 8-neighborhood of the node---using a 4-neighborhood would most
likely yield better results, but we do not want to encode too much
task-knowledge into our model.
Using the one-hot encoding of the input, this leads to $9 \dot 5 = 45$ unary features.
With 11 output classes, the unary potential has $11 \cdot 9 \cdot 5=495$ parameters.
Features for the pairwise potentials are constructed by concatenating the features
of the two neighboring nodes, taking the direction of the edge into account.
The pairwise feature therefore has dimensionality $45 \cdot 4$, with the first
$45$ entries corresponding to the feature of the ``top'' node, the second $45$
entries to the features ``bottom'' node, followed bye the ``left'' and
``right'' nodes. As each edge is either horizontal \emph{or} vertical, only two
of these parts will be non-zero for any given edge.  With $45 \cdot 4$ edge
features, the pairwise potentials have $45 \cdot 4 \cdot 11^2 =21780$ parameters.

\subsection{Superpixel CRFs for Semantic Segmentation}
Our main attention is devoted to the use of conditional random fields for
semantic segmentation.  We use the Pascal VOC and MSRC datasets.
The Pascal VOC has 964 training, each having
around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
each. There are 21 classes in the MSRC-21 dataset.
Each superpixel is represented as an output variable, with the ground truth obtained
by majority vote over the pixel belonging to the superpixel. We removed all
superpixels which were mostly ``void'', leading to some samples haven much
fewer than 100 variables.
Pairwise potentials were introduced for each pair of neighboring superpixels.
We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC21 dataset.
We use the same three pairwise features for both dataset: a constant, a color
feature and a feature encoding relative vertical position.
Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
2646$ parameters for the potentials on the MSRC21 dataset.

%For Pascal VOC we use the probabilities provided by
%\citet{krahenbuhl2012efficient}, which combine low-level Texton-Boost
%potentials with bounding-box object detectors.  We average probabilities inside
%each superpixel, yielding 21 input features for both datasets.  We also used
%the
%same pairwise features for the two datasets: a constant feature, a color
%contrast feature, and a feature encoding relative vertical position between
%superpixels.  Overall, the models for both datasets had $21 \cdot 21 + 21^2
%\cdot 4$ features.  It is possible to assume symmetric potentials for this
%dataset, leading to less pairwise features. We found both parametrizations
%yielded similar results.

\begin{table}
    \begin{tabu} to \linewidth{@{}X[1,l]YX[1,c]X[1,c]YY@{}}
    %\multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\
    \toprule
    Dataset    & {samples} & variables & graph & {dim $\theta$} & {labels}\\
    \cmidrule{1-6}
    MNIST      & 60000   & 1     & none  & 7840  & 10\\
    OCR-small  & 704     & 3--14 & chain & 1444  & 26\\
    OCR-large  & 6173    & 3--14 & chain & 1444  & 26\\
    scene-tree & 1211    & 6     & tree  & 1244  & 2\\
    scene-full & 1211    & 6     & loopy & 1284  & 2\\
    yeast-tree & 1500    & 14    & tree  & 1494  & 2\\
    yeast-full & 1500    & 14    & loopy & 1806  & 2\\
    snake      & 200     &84--168& loopy & 22275 & 11\\
    MSRC21     & 276     & 7--113& loopy &  1764 & 21\\
    Pascal VOC & 964     & 8--112& loopy &  2646 & 21\\
    \bottomrule
    \end{tabu}
    \caption{Summary of datasets used in the evaluation. \tablabel{datasets}}
\end{table}

\section{Experiments}
We compare the following algorithms on the above models:
\begin{itemize}
    \item Stochastic Subgradient Descent (SSGD) using the Pegasos schedule for step-sizes.
    %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
    \item The $1$-slack cutting plane algorithm without constraint caching.
    \item The $1$-slack cutting plane algorithm with constraint caching. For
            each sample, the last $50$ inference results are stored.
            Details of the caching strategy can be found in Chapter~\ref{ch:exact_learning}.
            %TODO
    \item The $n$-slack cutting plane algorithm, where the QP is solved after each training.
    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 training samples.
    \item The SZLJSP algorithm.
\end{itemize}

All algorithm take a $C$ parameter, which we adjusted on a fully trained model
on a validation set (experiments not reported here) and held constant for all
models.  We found, however, that the algorithms and models are quite robust to
the choice of $C$ within one or two orders of magnitude.
As stopping criterion, we used a duality gap of $0.1$ when possible, and a
pre-defined number of iterations for the subgradient algorithms. It is worth
noting that the quadratic programming based algorithms have additional
hyper-parameters that we do not discuss here in detail, such as the threshold
for removing a constraint as inactive, how often inactive constraints are
removed, and parameters of the underlying QP solver. On the other hand, the
SZLJSP, for example, has no hyper-parameters except for the stopping criterion.

Our ultimate evaluation criterion is how fast the learning algorithms converge,
and how robust they are to approximate inference.  To quantify our goals, we
track primal suboptimality and training set error during learning, as well as
final error on the test set.
We report suboptimality as a function of runtime, and as a function of passes
over the training set.  Both are important to practitioners, as they give
important insight into the working of the algorithm.  The actual runtime is
certainly the most important factor, but also highly influenced by
implementation details and properties of the dataset. We are particularly
interested in cases where inference is highly non-trivial, making inference the
dominating factor of runtime.
We define one pass over the dataset as calling prediction once for each sample.
This way, we can get a clear picture how learning times with complexity of the
inference task. For the subgradient and SZLJSP algorithms, learning time is linear
in the number of passes over the dataset, while caching and solving of a QP makes
the learning time depend on the number of iterations in hard-to-predict ways.
All algorithms were run on a single core i7 processor.

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these
tasks, exact inference is possible, and we can compute the exact objective of
the various algorithms easily. We use the dynamic programming algorithm
implemented in \textsc{OpenGM} for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
We found that using the linear programming relaxation on the multi-label
dataset is often tight, and there was no need to resort to approximate
inference.

We start our experiments with the MNIST dataset.
The convergence of the primal objective and training set loss in terms of
passes over the training set is shown in \Figref{mnist_iterations}.
Several trends can be observed: the $n$-slack cutting plane algorithms
converge fastest, with the version that recomputes the QP at every step leading
the race. The algorithms are closely followed by the caching $1$-slack cutting
plane algorithm.  They are followed by the significantly slower SZLJSP
algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
subgradient descent.  These results are intuitive, as they reflect ``how much
work'' each algorithm does for each loss-augmented prediction step. More work
towards the objective leads to faster convergence.  This ``more work'' is
quantified in \Figref{mnist_times}.  Clearly the $1$-slack algorithms to ``to
much'' work, leading to very slow convergence.  Also, caching does not speed up
learning learning on this dataset, as loss-augmented prediction is as fast as
querying the cache.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on OCR-small. 
\figlabel{ocr_small_results}}
\end{figure}

We now consider datasets with non-trivial inference.
The results for OCR-small are shown in \Figref{ocr_small_results}.
The trends seen in the figures are similar to what we saw for MNIST above.
Again, even though inference is non-trivial, the non-caching $1$-slack
algorithm is slower than the caching one. Again, the plots against passes over the training set
do not reflect the runtime complexity of the $n$-slack approach.

\begin{figure}
    \begin{center}
    \includegraphics[width=.5\linewidth]{evaluation/images/letters_small_dual}
\end{center}
\caption{%
   Convergence of the primal objective on OCR-small. 
\figlabel{ocr_small_dual}}
\end{figure}

There is one interesting phenomenon to be seen here: The cutting-plane algorithms stop immediately
when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
on learning much longer---even longer than it took to get to the desired suboptimality. This is caused
by a much looser bound given by the dual. Remember that the stopping criterion for both cutting-plane 
and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
This means that in for this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
in guaranteeing the desired suboptimality.
\Figref{ocr_small_dual} illustrates this point by ploting primal and dual
objectives for the $1$-slack cutting plane and the SZLJSP algorithms.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on OCR-large. 
\figlabel{ocr_large_results}}
\end{figure}

The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
algorithm when solving the QP at every step here.
The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
with respect to runtime. The SZLJSP algorithm converges fast intially, but the $1$-slack cutting plane
is faster at high precision and faster in certifying the desired duality gap.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on the multi-label datasets. From top to
   bottom: the scene dataset using a tree model, the scene dataset using a full
   model, the yeast dataset using a tree model, and the yeast dataset using a
   full model.
\figlabel{multi-label-results}}
\end{figure}

Next, we consider the multi-label task, starting with the tree-shaped models.
The results for the yeast and scene datasets are shown in \Figref{multi-label-results},
respectively. The caching $1$-slack algorithm is much faster than the
non-caching one on both datasets. While the two variants of the $n$-slack algorithm are equally
fast with respect to passes over the training data, solving the QP every 100 steps is among the
fastest algorithms, while solving the QP every step is among the slowest.

The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
lag behind the stochastic algorithms on the larger yeast dataset. Again, initial convergence is
faster for the stochastic algorithms.

%, inference is much more complex than in the previous tasks.

\subsection{Experiments using Approximate Inference}
For the remaining tasks of multi-label classification with the full graph, the
snakes dataset and image segmentation, exact inference is usually not possible.

We use two different approximate inference algorithms, fusion moves, a fast
local search procedure, and AD$^3$, a dual-subgradient method to solve the LP
relaxation of the inference problem.
Evaluation of algorithms where exact inference is not possible is much harder,
as it is usually not possible to evaluate the exact objective. When using AD$^3$,
we can still consider the relaxed task, which will be solved exactly in the
majority of cases (AD$^3$ can fail to find the primal solution to the LP
relaxation).
When using fusion moves, there is no obvious interpretation of the produced
primal objective, other than that it provides a lower bound of the actual
objective. Nevertheless, we find it informative to analyze the behavior of this
lower bound.

For both inference algorithms, we also evaluate the predictive performance on
the training set.  We explicitly do not consider the relaxed problem
here---instead we round possible fractional results. While the training error
does not directly reflect the objective, it provides a measure of how effective
the ``prediction machine'' given by the learned model together with the
inference algorithm is.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}
    \end{tabu}
\caption{%
    Results on the snakes dataset. The top four plots show results using fusion
    move inference, the bottom four plots using AD$^3$. The training set error shows
    that none of the algorithms is able to solve the task using fusion moves.
    The objective values for fusion move inference are only approximate, giving
    an upper bound on the true primal objective, while the objective for AD$^3$
    is the true objective for the relaxed task.
\figlabel{snakes_results}}
\end{figure}

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time_loss}
    \end{tabu}
\caption{%
    Objective and traning set loss on the segmentation task. The top four plots
    show results for the MSRC-21 dataset, the bottom four plots for the Pascal
    VOC dataset. As approximate inference was used, the shown objective values
    are only lower bounds on the actual primal objective.
\figlabel{msrc_pascal}}
\end{figure}

\section{Summary}
From the experiments above, it is clear there is currently \emph{no single best algorithm}.
However, there are some clear trends:
\begin{itemize}
    \item SZJLSP always converges faster than SSGD, with each iteration having
        the same time complexity.  This indicates that SZJLSP should be
        preferred over SSGD, apart from settings with strong memory
        constraints.
    \item In nearly all experiments, caching significantly improved performance
        of the 1-slack cutting plane algorithm.
    \item Initially SZJLSP and even SSGD converge faster than the 1-slack
        cutting plane algorithm, while the 1-slack cutting plane algorithm was
        faster in achieving high precision solutions and guaranteeing low
        duality gap.
    \item The n-slack cutting plane algorithm was often fastest to converge in
        terms of passes over the training set, but often slow in terms of
        runtime. The caching $1$-slack algorithm converged faster in terms of
        cpu-time on almost all datasets.
\end{itemize}
There are however some caveats with our analysis, in particular with respect to
the cutting plane solvers. The performance of the cutting plane algorithms
depends not only on the performance of the QP solver used (in our case
quadratic cone programming), but also on heuristics for pruning variables from
the QP, caching constraints (for the $1$-slack algorithm) and deciding how
often to solve the QP (for the $n$-slack algorithm). Better heuristics could
have a beneficial effect on these algorithms. On the other hand, it is
remarkable how competative the SZLJSP algorithm is, without the need for any
such implementation tricks. This makes the SZJLSP algorithm very attractive for
the practitioner who does not want to spend much time tuning the
implementation.

We do not consider SSGD using averaging here, but found it so produce similar
results in tentative experiments. In particular, it was never faster than SZLJSP.

With respect to approximate learning, we found that the inference algorithm has
a stronger impact on the performance than the learning algorithm.
While the SSGD and SZJLSP algorithms have somewhat lower training error on the snakes than
the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
task near perfect.
For the segmentation datasets, the approximate inference seem to be less of a
problem, as most algorithms achieve the desired approximate duality gap (which
is not the actual duality gap, see Chapter~\ref{ch:exact_learning} for a
discussion).
The large label space made using the linear programming relaxation intractable,
though, and we can not draw a final conclusion here. In the next chapter,
Chapter~\ref{ch:exact_learning}, we will demonstrate how we can learn exactly,
even in this setting.